% Chapter I: Preliminaries: Set theory and categories
\chapter{Preliminaries: Set theory and categories}

\section{Naive set theory}
\begin{xca}
Let $U = \{ x \mid x \not\in x \}$. Then, $U \not\in U \iff U \in U$, a
contradiction. This is Russell's paradox. Either we assume the \emph{set of all
sets} doesn't exist, or we need to give up the axiom of \emph{unrestricted
comprehension} in set theory.
\end{xca}

\begin{xca}
Suppose $\sim$ is an equivalence relation on a set $S$. For every element $a \in
S$, define the \emph{equivalence class} of $a$ (with respect to $\sim$) by
\begin{center}
$[a]_{\sim} := \{ b \in S \mid b \sim a \}$.
\end{center}
Then, we note that due to \emph{reflexivity}, the equivalence class
$[a]_{\sim}$ of every element $a \in S$ contains $a$, and hence, is nonempty.
Also, $[a]_{\sim} \subset S$, and therefore, $\bigcup_{a \in S} [a]_{\sim} =
S$. Finally, we show the equivalence classes are mutually disjoint. Indeed, for
any two elements $a, b \in S$, if $[a]_{\sim}$ and $[b]_{\sim}$ are disjoint,
then there is nothing to prove. So, suppose $[a]_{\sim} \cap [b]_{\sim}$ is
nonempty. Then, there exists some $c \in S$ that belongs to such an
intersection. Thus, $c \sim a$ and $c \sim b$. By symmetry, $a \sim c$, and
thus, by transitivity, $a \sim b$, which by symmetry again, implies $b \sim a$.
Therefore, for all $x \in [a]_{\sim}$, we have $x \sim a$, and since $a \sim
b$, by transitivity, $x \sim b$, which implies $x \in [b]_{\sim}$, from which
we conclude $[a]_{\sim} \subset [b]_{\sim}$. We can similarly show $[b]_{\sim}
\subset [a]_{\sim}$. Hence, $[a]_{\sim} = [b]_{\sim}$. This establishes
equivalence classes are mutually disjoint. Hence, the set $\mathscr{P}_{\sim}$
of equivalence classes of $S$ is indeed a partition of $S$.
\end{xca}

\begin{xca}
Suppose $\mathscr{P}$ is a partition on a set $S$. Define a relation $\sim$ on
$S$ as follows: For any two elements $a, b \in S$, $a \sim b$ iff $a$ and $b$
belong to the same set in the partition. Then, it is easy to check $\sim$ is
indeed an equivalence relation on $S$. $\mathscr{P}$ is, therefore, the
corresponding partition of the aforesaid equivalence relation, and we are done.
\end{xca}

\begin{xca}
Note the set of equivalence relations on a set $S$ are in a one-to-one
correspondence with the set of partitions of $S$. Thus, the number of different
equivalence relations that may be defined on $S = \{ 1, 2, 3\}$ equals the
number of partitions of $S$, and this number equals $5$, since the partitions
of $S$ are
\begin{center}
    $\{ \{ 1 \}, \{ 2 \}, \{ 3 \} \}, \{ \{ 1, 2 \}, \{ 3 \} \}, \{ \{ 1, 3 \},
    \{ 2 \} \}, \{ \{ 2, 3 \}, \{ 1 \} \}, \{ \{ 1, 2, 3 \} \}$.
\end{center}
The above partitions are also written as $1|2|3, 12|3, 13|2, 23|1, 123$.
\end{xca}

\begin{xca}
An example of a relation $R$ (defined on a set $S$) that is reflexive and
symmetric but not transitive is the following:
\begin{center}
$R = \{ (1,1), (2,2), (3,3), (1,2), (2,1), (2,3), (3,2) \}$, where
$S = \{ 1,2,3 \}$.
\end{center}
\end{xca}

\begin{xca}
Define a relation $\sim$ on the set $\R$ of real numbers by setting
\begin{center}
 $a \sim b \iff b - a \in \mathbb{Z}$.
\end{center}
We claim $\sim$ is an equivalence relation. To that end, note, for all $a \in
\R$, we have $a \sim a$, since $a - a = 0 \in \Z$. Therefore, $\sim$ is
reflexive. Also, if $a \sim b$, then $b - a \in \Z$, which implies $a - b \in
\Z$, and thus, $b \sim a$. Therefore, $\sim$ is symmetric. Finally, suppose $a
\sim b$ and $b \sim c$. Then, $b - a, c - b \in \Z$, and thus, $c - a = (c - b)
+ (b - a) \in \Z$. Thus, $\sim$ is transitive. Therefore, $\sim$ is an
equivalence relation on $\R$.

(Description of $\sim$) Note all reals that have the same decimal expansion
belong to the same equivalence class under $\sim$. Thus, $[0]_{\sim} = \Z$, and
for any $0 < \alpha < 1, [\alpha]_{\sim} = \{ n + \alpha \mid n \in \Z \}$. This
takes care of all the reals, since each real can always be written as $n +
\alpha$, for some $n \in \Z$ and $0 < \alpha < 1$. Therefore, a `compelling'
description for $\R/{\sim}$ is the unit interval $[0,1]$, such that the
endpoints, $0$ and $1$, are `glued' together. In other words, it is a `loop'
or a 1-sphere.\\

Define a relation $\approx$ on the plane $\R \times \R$ as follows:
\begin{center}
$(a_1, a_2) \approx (b_1, b_2) \iff b_1 - a_1 \in \Z \text{ and }
b_2 - a_2 \in \Z$.
\end{center}
Then, just as above, it is easy to show $\approx$ defines an equivalence
relation on $\R \times \R$. We note $[(0,0)]_{\sim} = \{ (m,n) \mid m,n \in \Z
\}$, and for any $0 < \alpha, \beta < 1, [(\alpha, \beta)]_{\sim} = \{ (m +
\alpha, n + \beta) \mid m,n \in \Z \}$. Thus, a `compelling' description of $\R
\times \R/{\approx}$ is the unit square $[0,1] \times [0,1]$ with the four
corners joined together, so that it forms a 2-sphere.
\end{xca}

\section{Functions between sets}
\begin{xca}
We claim the number of bijections from a set $S$ with $n$ elements to itself is
$n!$. To begin with, any element in $S$ can be mapped to any of the $n$ possible
elements in $S$. Then, the next element in $S$ can be mapped to any of the
remaining $n-1$ elements in $S$, and so on, with the last element in $S$ being
mapped to the last remaining element in $S$. Thus, the number of bijections
equals $n \cdot (n-1) \cdot \ldots \cdot 1 = n!$, which proves our claim.
\end{xca}

\begin{xca}
Assume $A \neq \emptyset$, and let $f: A \to B$ be a function. We claim $f$ has
a right inverse iff it is surjective.

($\impliedby$) Suppose $f$ has a right inverse, $g: B \to A$, say. Then,
$f \circ g = 1_B$. Thus, for all $b \in B$, $b = 1_B(b) = (f \circ g)(b) =
f(g(b)) = f(a)$, where $g(b) = a \in A$. This shows $f$ is surjective.

($\implies$) Suppose $f$ is surjective. Then, for any $b \in B$, the fiber of
$f$ over $b$ is nonempty. Thus, $\{ f^{-1}(b) \}_{b \in B}$ is a family of
nonempty sets, and therefore, using the \emph{axiom of choice}, we can construct
a function $g : B \to A$ as follows: For all $b \in B$, $g(b) = a$ for some
$a \in f^{-1}(b)$. Hence, for all $b \in B, (f \circ g) (b) = f(g(b)) = f(a) =
b = 1_{B} (b)$, and so, $f \circ g = 1_B$. This establishes $g$ is the right
inverse of $f$, and we are done.
\end{xca}

\begin{xca}
Suppose $f : A \to B$ is a bijection. Then, $f$ has an inverse $f^{-1}: B \to A$
such that $f^{-1} \circ f = 1_A$ and $f \circ f^{-1} = 1_B$. Clearly, $f$ is an
inverse of $f^{-1}$, showing $f^{-1}$ is also a bijection.

Suppose $f: A \to B$ and $g: B \to C$ are bijections. We claim $g \circ f: A \to
C$ is also a bijection. To that end, we show $f^{-1} \circ g^{-1}: C \to A$ is
the inverse of $g \circ f$. Indeed, $(g \circ f) \circ (f^{-1} \circ g^{-1}) =
g \circ (f \circ f^{-1}) \circ g^{-1} = g \circ 1_B \circ g^{-1} = g \circ
g^{-1} = 1_C$. And, $(f^{-1} \circ g^{-1}) \circ (g \circ f) = f^{-1} \circ
(g^{-1} \circ g) \circ f = f^{-1} \circ 1_B \circ f = f^{-1} \circ f = 1_A$,
and we are done.
\end{xca}

\begin{xca}
We show `isomorphism' is an equivalence relation on any set of sets.\\
(Reflexivity) For all sets $A, 1_A : A \to A$ is a natural bijection, and thus,
$A \cong A$.\\
(Symmetry) Suppose $A \cong B$ for any two sets $A, B$. Then,
there exists a bijection $f : A \to B$, such that its inverse $f^{-1} : B \to
A$ is also a bijection (as shown in the above exercise.) Thus, $B \cong A$.\\
(Transitivity) Finally, suppose for any three sets, $A, B$ and $C$, $A \cong B$
and $B \cong C$, with $f : A \to B$ and $g: B \to C$ as bijections. Then, from
the previous exercise, $g \circ f : A \to C$ is also a bijection, and thus $A
\cong C$.\\
Thus, our original claim is established.
\end{xca}

\begin{xca}
(\emph{\textbf{Epimorphism}}) A function $f : A \to B$ is an \emph{epimorphism}
(or \emph{epi}) if the following holds: For all sets $Z$ and all functions
$\alpha ', \alpha '' : B \to Z$,
\begin{center}
$\alpha ' \circ f = \alpha '' \circ f \Longrightarrow \alpha ' = \alpha ''$.
\end{center}
In other words, an epimorphism $f$ is \emph{right cancellative}.\\

\emph{Proposition}: A function is surjective iff it is an epimorphism.

\emph{Proof}. ($\implies$) Suppose $f: A \to B$ is an epimorphism. Assume, for
the sake of contradiction, $f$ is \emph{not} surjective. Then, there exists an
element $b_0 \in B$, such that, for all $a \in A$, $f(a) \neq b_0$. We now
construct two distinct functions $\alpha', \alpha'' : B \to \{ 0, 1 \}$ as
follows:
\begin{center}
\[
\alpha'(b) = 0
\]

\[
\alpha''(b) =
    \begin{cases}
        0 & \text{if } b \neq b_0 \\
        1 & \text{if } b = b_0
    \end{cases}
\]
\end{center}
Then, it is easy to check that, for all $a \in A$,
$(\alpha'' \circ f)(a) = \alpha''(f(a)) = 0 = \alpha'(f(a)) = (\alpha' \circ
f)(a)$, which implies $\alpha' \circ f = \alpha'' \circ f$. However, $\alpha'
\neq \alpha''$, which contradicts our assumption that $f$ is an epimorphism.
Hence, we conclude $f$ is surjective.

($\impliedby$) Suppose $f: A \to B$ is surjective. Then, it has a right inverse
$g: B \to A$ such that $f \circ g = 1_B$. Now, assume, for any set $Z$ and any
two functions $\alpha', \alpha'': B \to Z$, $\alpha' \circ f = \alpha'' \circ
f$. Then, $\alpha' = \alpha' \circ 1_B = \alpha' \circ (f \circ g) = (\alpha'
\circ f) \circ g = (\alpha'' \circ f) \circ g = \alpha'' \circ (f \circ g) =
\alpha'' \circ 1_B = \alpha''$, thus proving $f$ is an epimorphism.
\end{xca}

\begin{xca}
Any function $f: A \to B$ determines a section $g: A \to A \times B$ of $\pi_A:
A \times B \to A$ by defining $g$ as follows:
\begin{center}
    $a \mapsto (a, f(a))$
\end{center}
Then, for all $a \in A$, $(\pi_A \circ g)(a) = \pi_A(g(a)) = \pi_A(a, f(a)) =
a = 1_A(a)$, which implies $\pi_A \circ g = 1_A$, thereby showing $g$ as
defined above is indeed a section of $\pi_A$.
\end{xca}

\begin{xca}
Let $f: A \to B$ by any function. We show the graph $\Gamma_f$ of $f$ is
isomorphic to $A$. First, recall the definition of $\Gamma_f$:
\begin{center}
    $\Gamma_f := \{ (a, b) \in (A, B) \mid b = f(a) \} \subseteq A \times B$.
\end{center}
We define a function $g: A \to \Gamma_f$ by
\begin{center}
    $a \mapsto (a, f(a))$.
\end{center}
Then, for any $(a, b) \in \Gamma_f$, we have $b = f(a)$, which implies $g(a) =
(a, f(a)) = (a, b)$, proving $g$ is surjective. Next, for any $a', a'' \in A$,
suppose $g(a') = g(a'')$. This implies $(a', f(a')) = (a'', f(a''))$, which
implies $a' = a''$, thus proving $g$ is injective. Hence, $g$ is an isomorphism,
and so, $A \cong \Gamma_f$.
\end{xca}

\begin{xca}
We describe below explicitly all the terms in the canonical decomposition of the
function $f: \R \to \C$ defined by
\begin{center}
    $r \mapsto e^{2\pi ir}$.
\end{center}
Note $f$ determines an equivalence relation $\sim$ on $\R$ as follows: For all
$r', r'' \in \R$,
\begin{center}
    $r' \sim r'' \iff f(r') = f(r'')$.
\end{center}
Now, $f(r') = f(r'')$ whenever $e^{i2\pi r'} = e^{i2\pi r''}$, \emph{i.e.}
$e^{i2\pi(r' - r'')} = e^0$; that is, $2\pi(r' - r'') = 2\pi k$, where $k \in
\Z$. In other words, $f(r') = f(r'')$ iff $r' - r'' = k$, where $k \in \Z$.
This implies
\begin{center}
    $r' \sim r'' \iff r' - r'' \in \Z$.
\end{center}
The above equivalence relation matches the one stated in Exercise 1.6 of
Chapter 1. And, from the solution to the aforesaid exercise, we note that all
equivalence classes are of the form $[\alpha]_{\sim}$, where $\alpha \in [0,
1)$. Therefore, the isomorphism $\tilde{f}: \R/{\sim} \xrightarrow{\sim}
\opn{im} f$ is defined by
\begin{center}
    $\tilde{f}(\alpha) = e^{2\pi i\alpha}, \, \alpha \in [0, 1)$.
\end{center}
Also, note $\opn{im} f$ is the unit circle on the complex plane. So, the entire
decomposition of $f: \R \to \C$ is as shown below:
\[
\begin{tikzcd}
    \R \arrow[r, two heads]
       \arrow[rrr, bend left, "f"]
       & \R/{\sim} \arrow[r, "\sim"]{r}[swap]{\tilde{f}}
                   & \opn{im} f \arrow[r, hook]
                                & \C
\end{tikzcd},
\]
where $\R/{\sim}$ is a `loop' (1-sphere.)
\end{xca}

\begin{xca}
Suppose $A' \cong A''$ and $B' \cong B''$, and further $A' \cap B' = \emptyset$
and $A'' \cap B'' = \emptyset$. Then, there exist isomorphisms $f: A' \to A''$
and $g: B' \to B''$. Define a piecewise function $h: A' \cup B' \to A'' \cup
 B''$ as follows:
\[
    h(x) =
    \begin{cases}
        f(x) & \text{if } x \in A' \\
        g(x) & \text{if } x \in B'
    \end{cases}
\]
We claim $h$ is an isomorphism. Indeed, let $y \in A'' \cup B''$. Since $A''$
and $B''$ are disjoint, there are two cases to consider. Either, $y \in A''$,
in which case, since $f$ is surjective there exists some $x_1 \in A'$ such that
$f(x_1) = y$, which implies $h(x_1) = y$. Or, $y \in B''$, in which case, since
$g$ is surjective there exists some $x_2 \in B'$ such that $g(x_2) = y$, which
implies $h(x_2) = y$. Thus, in either case, $h$ maps some $x \in A' \cup B'$ to
$y \in A'' \cup B''$, thus showing $h$ is surjective.\\
Now, suppose $h(x_1) = h(x_2)$ for any $x_1, x_2 \in A' \cup B'$. There are
three cases to consider. First, if both $x_1, x_2 \in A'$, then $f(x_1) =
f(x_2)$, and since $f$ is injective, $x_1 = x_2$. Second, if both $x_1, x_2 \in
B'$, then $g(x_1) = g(x_2)$, and since $g$ is injective, $x_1 = x_2$. Finally,
without loss of generality, we can assume $x_1 \in A'$ and $x_2 \in B'$, which
implies $f(x_1) = g(x_2)$, but this is impossible, since $f(x_1) \in A''$ and
$f(x_2) \in B''$, but $A'', B''$ are disjoint. Thus, we conclude $h$ is
injective.\\
Hence, $h$ is an isomorphism.

The above demonstrates the \emph{disjoint union} operation $A \amalg B$ for any
two sets $A$ and $B$ is, indeed, well-defined \emph{up to isomorphism}.
\end{xca}

\begin{xca}
Suppose $A$ and $B$ are finite sets. Then, any one of the $|A|$ elements in $A$
can be mapped to any one of the $|B|$ elements in $B$. Thus, there are
exactly $|B|^{|A|}$ ways of constructing a function $A \to B$. Hence, $|B^A| =
|B|^{|A|}$.
\end{xca}

\begin{xca}
Let $A$ be a set, and let $2^A$ denote the set of functions from $A$ to $2 = \{
0, 1\}$. Then, define a mapping $f: 2^A \to \mc{P}A$, where $\mc{P}A$ is the
powerset of $A$, as follows: Any function $A \to 2$ is mapped to the subset
consisting of all those elements of $A$ that are mapped to $1$ under such a
function. Then, it is easy to check $f$ is both surjective and injective, and
hence, a bijection.
\end{xca}

\section{Categories}
\begin{xca}
Let $\mc{C}$ be a category. Consider a structure $\mc{C}^{op}$ with
\begin{itemize}
    \item $\opn{Obj}(\mc{C}^{op}) := \opn{Obj}(\mc{C})$;
    \item for $A, B$ objects of $\mc{C}^{op}$ (hence objects of $\mc{C}$),
    $\Hom_{\mc{C}^{op}}(A, B) := \Hom_{\mc{C}}(B, A)$.
\end{itemize}
We can make $\mc{C}^{op}$ into a category as follows.
\begin{itemize}
    \item For every object $A$ of $\mc{C}^{op}$, its identity morphism in
    $\mc{C}^{op}$ is defined as the identity morphism $1_A$ in $\mc{C}$.
    \item For any two morphisms $f \in \Hom_{\mc{C}^{op}}(A, B)$ and $g \in
    \Hom_{\mc{C}^{op}}(B, C)$, the composite $g \circ f \in
    \Hom_{\mc{C}^{op}}(A, C)$ is defined as the composite $f \circ g$ in
    $\mc{C}$.
    \item The associativity of composition in $\mc{C}^{op}$ is a direct
    consequence of associativity of composition in $\mc{C}$.
    \item The fact that identity morphisms in $\mc{C}^{op}$ are identities with
    respect to composition is a direct consequence of identity morphisms in
    $\mc{C}$ being identities with respect to composition.
\end{itemize}
The above construction, thus, shows $\mc{C}^{op}$ is a category. This
\emph{\textbf{opposite category}} is simply obtained by `reversing all the
arrows' in $\mc{C}$.
\end{xca}

\begin{xca}
If $A$ is a finite set, then $|\End_{\catname{Set}}(A)| = |A^A| = |A|^{|A|}$.
\end{xca}

\begin{xca}
In Example 3.3, to say $1_a = (a, a)$ is the identity of element $a \in S$ with
respect to composition means that any $f = (a, b) \in \Hom(a, b)$ remains
unchanged when composed with $1_a$ or $1_b$ in the appropriate manner. In the
first case, we have $a \sim a$ and $a \sim b$, and thus, by transitivity, $a
\sim b$, from which we conclude $f 1_a = (a, b) = f$. Similarly, in the second
case, we conclude $1_b f = f$.
\end{xca}

\begin{xca}
Suppose we can define a category in the style of Example 3.3 using the relation
$<$ on the set $\Z$. Then, any element (object) $n \in \Z$ will have an identity
morphism $1_n = (n, n)$ that would imply $n < n$, which is impossible. Hence,
we can't define such a category using the relation $<$ on the set $\Z$.
\end{xca}

\begin{xca}
Example 3.4 is an instance of the categories considered in Example 3.3 in the
sense that set inclusion $\subseteq$ (morphisms) between subsets of $S$
(objects) satisfies the reflexivity and transitivity axioms for $\sim$.
\end{xca}

\begin{xca}
(Assuming some familiarity with linear algebra.) Define a category $\mathsf{V}$
by taking $\opn{Obj}(\mathsf{V}) = \N$ and letting $\Hom_{\mathsf{V}}(n, m) =$
the set of $m \times n$ matrices with real entries, for all $n, m \in \N$.
Composition is defined as products of matrices. Note, for all $n \in \N$,
\[
\Hom_{\mathsf{V}}(n, 0) = \emptyset = \Hom_{\mathsf{V}}(0, n).
\]
This category does \emph{not} `feel' familiar, since it is quite unlike any of
the other categories we have seen before. Objects of $\mathsf{V}$ are not like
the usual sets or elements we have encountered so far, and its morphisms are
matrices which are not the typical set-theoretic functions.
\end{xca}

\begin{xca}
Let $\mc{C}$ be a category, and $A$ an object of $\mc{C}$. We define a category
$\mc{C}^A$ as follows:
\begin{itemize}
    \item $\opn{Obj}(\mc{C}^A) :=$ collection of all morphisms from $A$ to any
    object of $\mc{C}$; that is, an object of $\mc{C}^A$ is a morphism $f \in
    \Hom_{\mc{C}}(A, B)$ for some object $B$ of $\mc{C}$.
    \item For any two objects $f_1: A \to B$ and $f_2: A \to C$ of $\mc{C}^A$,
    morphisms $f_1 \to f_2$ are defined to be commutative diagrams
    \[
    \begin{tikzcd}
        & A \arrow[dl, swap, "f_1"]
            \arrow[dr, "f_2"]
            & \\
            B \arrow[rr, swap, "\sigma"]
              & & C
    \end{tikzcd}
    \]
    in the `ambient' category $\mc{C}$. That is, morphisms $f_1 \to f_2$
    correspond precisely to those morphisms $\sigma: B \to C$ in $\mc{C}$ such
    that $\sigma \circ f_1 = f_2$.
\end{itemize}

(Composition) Two morphisms $f_1 \to f_2$ and $f_2 \to f_3$ in $\mc{C}^A$
correspond to putting two commutative diagrams side-by-side:
\[
\begin{tikzcd}
    & A \arrow[dl, swap, "f_1"]
        \arrow[d, swap, "f_2"]
        \arrow[dr, "f_3"]
        & \\
    B \arrow[r, swap, "\sigma"]
      & C \arrow[r, swap, "\tau"]
          & D
\end{tikzcd}
\]
Then, the diagram obtained by removing the central arrow, \emph{i.e.}
\[
\begin{tikzcd}
    & A \arrow[dl, swap, "f_1"]
        \arrow[dr, "f_3"]
        & \\
    B \arrow[rr, swap, "\tau \circ \sigma"]
      & & D
\end{tikzcd}
\]
also commutes, since $\mc{C}$ is a category.
\end{xca}

\begin{xca}
A \emph{subcategory} $\mc{C}'$ of a category $\mc{C}$ consists of a collection
of objects of $\mc{C}$, with morphisms $\Hom_{\mc{C}'}(A, B) \subseteq
\Hom_{\mc{C}}(A, B)$ for all objects $A, B$ in $\opn{Obj}(\mc{C}')$, such that
identities and compositions in $\mc{C}$ make $\mc{C}'$ into a category. A
subcategory $\mc{C}'$ is \emph{full} if $\Hom_{\mc{C}'}(A, B) =
\Hom_{\mc{C}}(A, B)$ for all $A, B$ in $\opn{Obj}(\mc{C}')$.

We can construct a category of \emph{infinite sets} as follows. The objects of
this category are all infinite sets from $\catname{Set}$, and morphisms from an
infinite set $A$ to another infinite set $B$ are just the usual set-functions.
Then, it is clear that this category is indeed a full subcategory of
$\catname{Set}$.
\end{xca}

\begin{xca}
% TODO: Chapter I - Exercise 3.9
\end{xca}

\begin{xca}
Given any set $A$ in $\catname{Set}$, the subsets (subobjects) of $A$ are in a
one-to-one correspondence with functions $A \to \{ 0, 1\}$. Thus, $\Omega =
\{ 0, 1 \}$ is a \emph{subobject classifier} of $\catname{Set}$.
\end{xca}

\begin{xca}
The category $\mc{C}^{A, B}$ consists of the following pieces of data:
\begin{itemize}
    \item $\opn{Obj}(\mc{C}^{A, B})$ = diagrams
    \[
    \begin{tikzcd}
        A \arrow[rd, "f"] & \\
        & Z \\
        B \arrow[ru, swap, "g"]
    \end{tikzcd}
    \]
    \item morphisms $(f_1: A \to Z_1, g_1: B \to Z_1) \to (f_2: A \to Z_2,
    g_2: B \to Z_2)$ are commutative diagrams
    \[
    \begin{tikzcd}
        A \arrow[rd, "f_1"]
          \arrow[rrd, bend left, "f_2"]
          & & \\
        & Z_1 \arrow[r, swap, "\sigma"]
              & Z_2 \\
        B \arrow[ru, swap, "g_1"]
          \arrow[rru, swap, bend right, "g_2"]
          & &
    \end{tikzcd}
    \]
    \item (Composition) Given commutative diagrams
    \[
    \begin{tikzcd}
        A \arrow[rd, "f_1"]
          \arrow[rrd, bend left, "f_2"]
          \arrow[rrrd, bend left, "f_3"]
          & & & \\
        & Z_1 \arrow[r, swap, "\sigma"]
              & Z_2 \arrow[r, swap, "\tau"]
              & Z_3 \\
        B \arrow[ru, swap, "g_1"]
          \arrow[rru, swap, bend right, "g_2"]
          \arrow[rrru, swap, bend right, "g_3"]
          & & &
    \end{tikzcd},
    \]
    the following diagram commutes:
    \[
    \begin{tikzcd}
        A \arrow[rd, "f_1"]
        \arrow[rrd, bend left, "f_3"]
        & & \\
        & Z_1 \arrow[r, swap, "\tau \circ \sigma"]
        & Z_2 \\
        B \arrow[ru, swap, "g_1"]
        \arrow[rru, swap, bend right, "g_3"]
        & &
    \end{tikzcd}
    \]
    \item (Identities) The identity morphism of
    \[
    \begin{tikzcd}
        A \arrow[rd, "f"] & \\
          & Z \\
        B \arrow[ru, swap, "g"]
    \end{tikzcd}
    \]
    is $Z \xrightarrow{1_Z} Z$, thus making the following diagram commute:
    \[
    \begin{tikzcd}
        A \arrow[rd, "f_1"]
          \arrow[rrd, bend left, "f_1"]
          & & \\
        & Z_1 \arrow[r, swap, "1_Z"]
              & Z_1 \\
        B \arrow[ru, swap, "f_2"]
          \arrow[rru, swap, bend right, "f_2"]
          & &
        \end{tikzcd}
    \]
    \end{itemize}~\\

    Given a category $\mc{C}$ and two fixed morphisms $\alpha: C \to A$ and
    $\beta: C \to B$, the category $\mc{C}^{\alpha, \beta}$ consists of the
    following pieces of data:
    \begin{itemize}
        \item $\opn{Obj}(\mc{C}^{\alpha, \beta})$ = commutative diagrams
        \[
        \begin{tikzcd}
            & A \arrow[rd, "f"] & \\
            C \arrow[ru, "\alpha"]
              \arrow[rd, swap, "\beta"]
              & & Z \\
            & B \arrow[ru, swap, "g"] &
        \end{tikzcd}
        \]
        in $\mc{C}$.
        \item Morphisms correspond to commutative diagrams
        \[
        \begin{tikzcd}
            & A \arrow[rd, "f_1"]
                \arrow[rrd, bend left, "f_2"] & \\
            C \arrow[ru, "\alpha"]
              \arrow[rd, swap, "\beta"]
              & & Z_1 \arrow[r, swap, "\sigma"]
                      & Z_2 \\
            & B \arrow[ru, swap, "g1"]
                \arrow[rru, swap, bend right, "g_2"] &
        \end{tikzcd}
        \]
        \item Composition and identities are defined in the obvious way.
        \item The usual composition and identity laws for $\mc{C}^{\alpha,
        \beta}$ follow from the composition and identity laws in $\mc{C}$.
    \end{itemize}
\end{xca}

\section{Morphisms}
\begin{xca}
Suppose $A_0 \xrightarrow{f_1} A_1 \xrightarrow{f_2} A_2 \xrightarrow{f_3}
\ldots \xrightarrow{f_n} A_n$ is a sequence of $n > 0$ morphisms in a
category $\mc{C}$. We claim any choice of placement of parentheses for $f_n
f_{n-1} \ldots f_1$ equals
\[(\ldots ((f_n f_{n-1}) f_{n-2}) \ldots) f_1.\]

We use induction on $n$ to prove our claim.

(Base cases) We note $f_1 = (f_1)$. And, also $f_2 f_1 = (f_2) f_1$. So the
statement holds for $n = 1, 2$.

(Inductive case) Suppose the statement holds for some $n > 1$. Now, consider
the placement of parentheses for the expression $E = f_{n+1} f_n f_{n-1} \ldots
f_1$. Any such choice will correspond to a representation of $E$ in the form
$e e'$, where both $e$ and $e'$ contain at least one term. This also implies
both $e$ and $e'$ contain at most $n$ terms. So, assume $e$ contains the terms
$f_{n+1}, f_n, \ldots, f_i$ and $e'$ contains the terms $f_{i-1}, \ldots, f_1$
for some $1 < i \le n + 1$. Then, by the inductive hypothesis,
\begin{align*}
    e &= (((f_{n+1} f_n) \ldots) f_i) \\
    e' &= (f_{i-1}(\ldots (f_2 f_1)))
\end{align*}
Then, through repeated use of the associativity of morphisms under composition,
we have $E$ \\
\begin{tabular}{l l}
    &= $e e'$ \\
    &= $e (f_{i-1}(\ldots (f_2 f_1)))$ \\
    &= $((((e f_{i-1}) f_{i-2}) f_{i-3}) \ldots) f_1$ \\
    &= $(\ldots ((f_{n+1} f_n) f_{n-1}) \ldots) f_1$
\end{tabular}
~\\

So, the statement holds for $n+1$ as well, and hence, by induction, the
statement holds for all naturals $n > 0$. And, we are done.
\end{xca}

\begin{xca}
We claim if a set $S$ is endowed with an equivalence relation $\sim$, then the
corresponding category $\mc{C}$ is a groupoid. Indeed, if $a \to b$ is a
morphism in $\mc{C}$, then $a \sim b$, which (by symmetry) implies $b \sim a$.
Thus, $b \to a$, and since there is at most one morphism from any object
(element) to another, we must have
\begin{align*}
    a \to b \to a &= 1_a \\
    b \to a \to b &= 1_b
\end{align*}
Hence, $a \to b$ is an isomorphism, for all elements $a, b \in S$, and thus,
$\mc{C}$ is a groupoid.
\end{xca}

\begin{xca}
Let $A, B$ be objects of a category $\mc{C}$, and let $f \in \Hom_{\mc{C}}(A,
B)$ be a morphism.
\begin{itemize}
    \item Suppose $f$ has a right inverse, $g: B \to A$, say. Then, $f \circ g
    = 1_B$. Assume $Z$ is an object of $\mc{C}$ and $\beta, \beta': B \to Z$
    morphisms such that $\beta \circ f = \beta' \circ f$. Then, $\beta = \beta
    \circ 1_B = \beta \circ (f \circ g) = (\beta \circ f) \circ g = (\beta'
    \circ f) \circ g = \beta' \circ (f \circ g) = \beta' \circ 1_B = \beta'$,
    thus showing $f$ is an epimorphism.
    \item In $(\Z, \le)$, considered as a category, any morphism is an
    epimorphism, but the same does not have a right inverse. For example, $1 \le
    2$, \emph{i.e.} $1 \to 2$. But, if the latter were to have a right inverse,
    then we would have $2 \to 1$, which would then imply $2 \le 1$, clearly a
    contradiction.
\end{itemize}
\end{xca}

\begin{xca}
Suppose $A \xrightarrow{f} B$ and $B \xrightarrow{g} C$ are both monomorphisms
in a category $\mc{C}$. Let $Z$ be any object of $\mc{C}$, and let $\alpha',
\alpha'': Z \to A$ be morphisms such that
\begin{center}
    $(g \circ f) \circ \alpha' = (g \circ f) \circ \alpha''$.
\end{center}
Then, $g \circ (f \circ \alpha') = g \circ (f \circ \alpha'')$, which implies
$f \circ \alpha' = f \circ \alpha''$, which implies $\alpha' = \alpha''$. Hence,
$g \circ f$ is a monomorphism. This demonstrates the composition of two
monomorphisms is a monomorphism.

Now, it is easy to check identity morphisms in $\mc{C}$ are monomorphisms. Thus,
one can define a subcategory $\mc{C}_{\text{mono}}$ of a category $\mc{C}$ by
taking the same objects as in $\mc{C}$ and defining
$\Hom_{\mc{C}_{\text{mono}}}$ to be the subset of $\Hom_{\mc{C}}(A, B)$
consisting of monomorphisms, for all objects $A, B$ of $\mc{C}$.\\

(Epimorphisms) Suppose $A \xrightarrow{f} B$ and $B \xrightarrow{g} C$ are both
epimorphisms in a category $\mc{C}$. Let $Z$ be any object of $\mc{C}$, and let
$\beta', \beta'': C \to Z$ be morphisms such that
\begin{center}
    $\beta' \circ (g \circ f) = \beta'' \circ (g \circ f)$.
\end{center}
Then, $(\beta' \circ g) \circ f = (\beta'' \circ g) \circ f$, which implies
$\beta' \circ g = \beta'' \circ g$, which implies $\beta' = \beta''$, thus
showing $g \circ f$ is an epimorphism. In a manner similar to the one shown
above, one can define a subcategory $\mc{C}_{\text{epi}}$ of a category
$\mc{C}$.
\end{xca}
